Joint Likelihood Deconvolution of Astronomical Images in the Presence of Poisson Noise

### 2023-10-04 src/tex/ms.tex
[showyourwork] overleaf sync

    The standard baseline method for image deconvolution was proposed independently by \cite{Richardson1972} and \cite{Lucy1974} (we will further refer to this method as \enquote{\textit{RL}}). Their method involves the

    involves treating the forward convolution by the PSF in a Bayesian and solving the e

    The method involves estimating the original image by iteratively convolving the observed image with the PSF and dividing by the result of convolving the estimated image with the transposed PSF. The iterations continue until the estimated image converges to a satisfactory solution.

@@

    To address the question of when to stop the RL iteration \cite{Reeves1995} proposed to use cross validation. The authors of \citep{Bi1994} proposed a bootstrapped based assessment of the significance of reconstructed image as as a stopping rule for RL.

    Another approach to image deconvolution is joint likelihood RL, which has been shown to be effective in creating a merged image from multiple images \cite{Ingaramo2014}. In this approach, RL takes the best out of each image and merges them to create a more accurate representation of the object.

@@

     \cite{Selig2015} and \cite{Pumpe2018} approached the problem of deconvolution of Poisson noise data implicitly using the framework of information field theory. Deconvolution, Denoising and Decomposition (D3PO) and (D4PO).

    Recently, image deconvolution using physical assumptions and decomposition into point and diffuse flux in 3D/4D has gained attention. proposed priors based on physical assumptions, which have been shown to work well for large-scale emission. However, the smoothness assumption in these methods can lead to artifacts in the galactic plane.

    Other interesting approaches to the problem of deconvolution can of course be found in the field of computer vision. For example \cite{Zoran2011} build on very similar assumption of local correlation for natural images. However instead of assuming a fixed local correlation structure, such as smoothness defined by a Gaussian with a correlation size, they proposed to learn the local correlation structure on millions of patches extracted from a set of reference images. This approach can be seen as a form of transfer learning. For this they proposed to split training images from representative data sets into small patches of size 8x8~pixels and fit a 64 dimensional Gaussian Mixture Model (GMM) to the distribution of extracted patches. Each pixel of the patch  is thereby treated as an independent dimension in the model. In total they chose to use $k=200$ Gaussian components. To use the GMM in reconstruction they introduce the expected patch likelihood (EPLL). They showed that the patch based GMM prior led to much improved image reconstructions for multiple inverse problems such as denoising, deblurring and inpainting. \cite{Bouman2016} later adapted the patch prior successfully for the reconstruction of radio data from the Event Horizon Telescope (EHT).

@@

    \subsection{Some Remarks on Deep Learning Methods}
    In the past decade Machine learning (ML) especially deep convolutional neural networks (CNN) became the standard for deblurring and super-resolution of natural images \todo{reference review?}. However their success relies on the availability of massive sets of training data with typically low noise and corresponding ground truth to perform supervised learning. For a variety of reasons, which we will outline in the following, this approach is not ideal for astronomical x-ray and \gammaray data.

@@

    To account for the possibility of different datasets having a different number of pixels we divide by $N$ (\todo{Does this need more justification...is it reasonable?}).
    By summing over all observations we finally get the joint log-likelihood of measuring a set of counts images $\mathbf{D}$ given an unknown flux image $\mathbf{x}$:
    %
    \begin{equation}
        \label{eq:joint}
        \mathcal{L}\left( \mathbf{D} | \mathbf{x} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D_j} | \mathbf{x} \right)
    \end{equation}

@@

    \newpage
    \subsection{Patch Based Image Prior}
    In general it is difficult to capture global image statistics in a prior likelihood taking into account all pixel to pixel correlations. The size of the corresponding correlation matrix would grow with the number of pixels squared and thus quickly becomes computationally untraceable. However astronomical images typically only show local correlations between pixels. This is because the spatial scale of astrophysical processes in an image is always limited by the finite interaction time and the fact that the interaction cannot propagate faster than the speed of light (structure in the CMB might be an exception here). This means on smaller scales astronomical images often contain basic structures such as edges, corners, filaments or periodic patterns formed by local astrophysical processes. For this reason we adopt the idea of a patch based image prior approach as introduced by \cite{Zoran2011}. Given a pre-trained Gaussian Mixture Model (GMM) for the distribution of pixel values in patches, we can evaluate a patch based prior for the reconstructed image $\mathbf{x}$ and approximate the \aposteriori likelihood.

@@

    \cite{Zoran2011} found that the reconstruction quality could be improved by working with a grid of overlapping patches. Such a grid is illustrated in Figure~\ref{fig:patches}. However using a grid of overlapping patches makes the interpretation of the prior as likelihood more difficult, because pixels at the corners and edges of patches are counted multiple times. For this we introduce a weighting mask when evaluating the likelihood of the GMM. This mask weights the pixels of the patch by their overlap with the neighboring patch. Overlapping edges are assigned a weight of 1/2, while overlapping corners are assigned a weight of 1/4, when computing the likelihood for a patch as given by  Equation~\ref{eq:gmm}. Apart from this we keep the process of choosing the most likely component unmodified. However we note that in this case we neglected the correlation between neighbouring patches. As the choice of the GMM component per patch is discrete, a simultaneous gradient based optimization of all patches is not easily possible, because we cannot compute gradients across GMM components. This would require a continuous representation or interpolation between GMM components. We propose ideas to work around this assumption later in the outlook of this manuscript.

@@

    \subsection{Choice and Tuning of Hyperparameters}
    \subsubsection{Cross Validation}
    Cross validation (CV) is an established method for model selection in statistics and machine learning to prevent a model from overfitting to the training data. To perform CV the available data is slit into a \enquote{training} and \enquote{validation} dataset. While optimizing the log-likelihood value is monitored  The use of multiple dataset allows to use cross-validation to prevent overfitting. One of the standard techniques in machine learning...

    Cross validation is feasible once sufficient data is available, which is for example the 
    case for deep \chandra~observations.

    CV as stopping rule for RL was proposed by \cite{Reeves1995}...

    \subsubsection{Other Approaches}

    Self supervised Poisson loss? J-invariance has been shown to work, Grid Search? Calibration on similar data with ground truth?
    Ground truth based calibration: record prior vs data likelihood on simulated data...

@@

    % \subsubsection{Chandra SNRs}
    % \label{sssec:chandra-snrs}. Because of the limited statistics we did not use any oversampling but kept the pixel Astrophysical processes forming the morphological structures are typically self-similar \todo{provide physical argument...I guess it is just the distance distribution?}. Structures appear on multiple angular scales. This means the prior could even be learned from the data itself if sufficiently free of noise. For analysis of Galactic sources with complex morphology and typically radially symmetric features such as supernova remnants (SNRs) we learned a prior from images of the bright \chandra SNRs \textit{Cas A} and \textit{G292.0+01.8}. In total we extracted \todo{insert number} patches and learned a GMM with 128 components.

@@

    \begin{itemize}
        \item Results are available in \url{https://github.com/jolideco/jolideco-benchmarks}
        \item We tried the binary search tree as proposed in \cite{Parameswaran2018},
        but without success, most likely because the native \texttt{Pytorch} support
        for parallelization speeds up the standard GMM quite a bit...
        \item Could use a Neural Net for GMM component selection as proposed in \cite{Rosenbaum15}
        \item GPU support via \texttt{Pytorch}
    \end{itemize}

@@

    \begin{itemize}
        \item \textbf{Pylira}: first we use the \textit{LIRA} method \citep{Esch2004} via the \textit{Pylira} Python wrapper \citep{Donath2022}. We rely on the default prior definition and keep the background norm fixed. In total we run the the sampling process for 5000 iterations and use 500 iterations as a \endquote{burn-in} period to allow the MCMC chain to converge. Then we compute the mean of the posterior distribution ignoring the burn-in period.

@@

    On scenario A1 all methods show acceptable to excellent results for the isolated point sources in the upper right of the image. This demonstrates all methods take into account the PSF and background models correctly. The early stopped uniform prior still shows an extension for those cases, because the optimization process has not finished. Optimizing for 1000 iterations clearly improves the result and shows almost perfect point sources corresponding to the size of approximately one pixel. The \jolideco method with the Zoran-Weiss patch prior results in less sharp images, because the prior has been learned on natural images of landscapes and every day objects and scenes, which do not include point sources. The \jolideco method with the prior learned from GLEAM data, which also includes point sources, shows much improved results. However not as sharp as the results obtained with the uniform prior. In the lower left of the image the point sources become fainter and move closer, challenging all methods in succeeding to separate those as independent sources. The uniform prior optimized \endquote{to end} shows good results, still separating the sources. All other methods cannot fully resolve weak and close by point sources. \jolideco with th e

    For scenario B3 all methods recover the point sources well. However large difference can be observed in the reconstruction of the Gaussian shaped extended sources. As expected, the uniform prior assumption lets the emission decomose into many point sources, neglecting the spatial correlation between them. \textbf{Pylira} shows slightly improved results with the gaps between the bright emission peeks being more filled, however still lacks a convincing result of reconstructing the source as a single object. The \jolideco method with the patch priors shows the best results both visually as well as measured by the SSI. The GLEAM v0.1 patch prior reconstructs the sources almost perfectly, visually almost indistinguishable from the ground truth. The Zoran \& Weiss prior also shows good reconstruction of the extended sources, but less smooth and leaving a more visible extension in the surrounding point sources.

    Scenario C3 challenges the methods to reconstruct a weak extended source under four brighter point sources. While again all methods reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. \textit{Pylira} shows a similar tendency but already improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features in the same scenario are decomposed into a \enquote{thread} of point sources by the uniform prior and LIRA method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that is possibly encoded in the training data.

@@

    \section{Application Examples}
    \subsection{Deep \chandra Observation}

    \begin{figure*}
        \script{example-chandra.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               Result of the \jolideco method applied to \chandra data of the SNR \textit{1E~0102.2–7219}. The left large image shows the summed counts from 25 observations. The right large image  shows the result from the \jolideco deconvolution process. The flux is given as counts relative to the reference observation with ID=1308. The panels \textit{Zoom A}, \textit{Zoom B} and \textit{Zoom C} show a zoomed-in version of three example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
        \end{centering}
    \end{figure*}

    \subsubsection{Image Example}
    As a first application example for real data we choose a set of short exposure observations of the SNR \textit{1E~0102.2–7219} by the \chandra telescope and combine them into a single deep observation using \jolideco and the patch prior learned from JWST data. In total we chose 25 observations\footnote{Same list of observation IDs as for \url{https://chandra.si.edu/photo/2009/e0102/}} taken with  the ACIS instrument and no grating.

    We reduced the data using \texttt{ciao} v4.15 and a custom \texttt{snakemake} workflow (see also Section~\ref{sec:reproducibility} on reproducibility). For the counts image we selected the energy range from \qty[mode = text]{0.5}{keV} to  \qty[mode = text]{7}{keV} and a pixel size of $0.25$ of the native pixel size of \chandra. This corresponds to an absolute  pixel size of $\approx$~\qty[mode = text]{0.125}{$\arcsec$/pixel} or a sampling of $\approx2$ pixels for the $38\%$ containment radius of Chandra's PSF. \chandra's instrument response varies with the observation offset and energy \todo{reference}, so we simulated the PSF for each observation independently using the tool \texttt{marx} \citep{Davis2012} and the \texttt{simulate\_psf} script. As reference spectrum for the PSF simulation we assumed the model derived by \cite{Plucinsky2017}. We converted the spectral from XSPEC to sherpa script using the helper script \texttt{convert\_xspec\_script}. As reported in \cite{Xi2019} the effect of pileup, even for the brightest regions is small ($<5\%$), so we did not include the effect of pileup in our PSF simulations.

@@

    \textbf{Zoom B} shows the region around the central compact object of the SNR \citep{Vogt2018}. While the emission is barely visible in the counts image it clearly sticks out above the background in the image reconstructed by \jolideco. The remaining visible extension in the point source is encoded in the JWST patch prior, which also shows a finite extension for point sources. As shown in Section~\ref{subsec:methods} for scenario A, the apparent size of the point source becomes smaller if the source becomes brighter.

@@

    Checking the residual images per observation (not shown here), we noticed that the assumption of the time independence of the source as well the instrument response is not completely fulfilled in this case. We saw the detection efficiency of \chandra changed between the observations as well as that sub-structure in the remnant expands over time. Both effects are previously known and have been reported for example in \cite{Xi2019} and \cite{}. Despite the clear improvement in spatial resolution the proper motion of structures in the image might limit the quality of the result. The full analysis example is available at \url{https://github.com/jolideco/jolideco-chandra-examples}.

@@

    The profile computed from the images reconstructed with \jolideco has a much smaller noise level and the peak is much more pronounced. By fitting the profile with a univariate spline and finding its roots we measure the full width half maximum of the curves. We find $\sigma=$ and $\sigma=$. Assuming that the PSF and the shape of peak of the profile are approximately Gaussian a we expect the following relation between the width of the counts profile, the PSF and the flux profile:

    \begin{equation}
        \sigma_{\mathrm{Flux}} = \sqrt{\sigma_{\mathrm{Counts}}^2 - \sigma_{\mathrm{PSF}} ^2}
    \end{equation}

    With $\sigma_{\mathrm{PSF}} \approx 0.25 \mathrm{arcsec}$ we obtain ..., which is consistent with the values we measured for the profile.

@@

    The \texttt{fermitools} output an energy dependent counts, exposure and PSF cube. To further reduce the data to images we use the \textit{Gammapy} package v1.0 \citep{GammapyZenodov1.0.1, Donath2023}. To compute an image of the expected background emission we rely on the latest Galactic diffuse emission model \texttt{gll\_iem\_v07.fits} and also include nearby sources from the 3FHL catalog \citep{Ajello2017}. We first fit the model components to the data using \textit{Gammapy}, excluding a circular region of radius \qty[mode = text]{1.02}{$\deg$} centered on Vela Junior. Finally we reduce this model to an image by summing over the energy dimension. To compute the exposure and PSF image we assume a power law with index $\Gamma=1.77$ \citep{Ajello2017} as a spectral model for Vela Junior and integrate over the energy dimension of the cube while weighting with the spectral model. We chose the GMM patch prior learned from \chandra SNRs described in the Section~\ref{sssec:chandra-snrs}. Because of the limited statistics we did not use any oversampling but keep the pixel size of the counts image.

@@

    Where $\hat{N}$ is the weighted sum of counts from a Gaussian smoothing kernel. As a check we compute a map of residuals based on the predicted counts, the resulting distribution of residual significances per pixel is centered at $\mu=-0.001$ with a width of $\sigma=0.991$, indicating a close agreement with a unit Gaussian distribution. More details of the example \fermi analysis are available at \url{https://github.com/jolideco/jolideco-fermi-examples}

      \begin{figure*}
        \script{example-fermi-lat.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-fermi-lat.pdf}
            \caption{
                Illustration of the result of the \jolideco method applied to \fermi data of the supernova remnant \textit{RX J0852.0-4622} or \textit{Vela Junior}. The left image shows the counts above \qty[mode = text]{10}{GeV}. The different event classes are stacked into a single image. The image in the center shows the flux reconstructed by the \jolideco method. The image on the right shows the residuals as computed by Equation~\ref{eq:approx-sigma} and smoothed with a \textit{Gaussian} kernel of width 5~pixels. More information on this analysis example can be found on \url{https://github.com/jolideco/jolideco-fermi-examples}. \todo{Move color bar to top and make images larger}
            }
            \label{fig:example-fermi-lat}
        \end{centering}
    \end{figure*}

>>>>

    The standard baseline method for image deconvolution in the presence of Poisson noise was proposed independently by \cite{Richardson1972} and \cite{Lucy1974} (we will further refer to this method as \enquote{\textit{RL}}). The method involves estimating the reconstructed image by an iterative procedure that converges to a maximum likelihood solution of the problem.

@@

@@

    An interesting modification of the standard RL method has been proposed by \cite{Ingaramo2014}. They showed that RL can be effective in creating a merged image from multiple observations. They also observed that RL would take the best visual properties out of an each image and merged them to create a more accurate estimate of the ground truth.

    \cite{Selig2015} and \cite{Pumpe2018} approached the problem of deconvolution of Poisson noise data implicitly using the framework of information field theory. Based on physical prior assumptions on the spatial distribution of point sources and local correlation structures of diffuse emission, they were able to deconvolve an astrophysical observation as well as decomposing it into point-like and diffuse emission components. While the method has worked well for large-scale diffuse emission, such as the Galactic diffuse gamma-ray emission, it seemed to fail for smaller extended emission structures such as mrophological details of extended sources.

    Other interesting approaches to the problem of deconvolution can of course be found in the field of computer vision. For example \cite{Zoran2011} build on very similar assumption of local correlation for natural images. However instead of assuming a fixed local correlation structure, such as smoothness defined by a Gaussian with a correlation size, they proposed to learn the local correlation structure on millions of patches extracted from a set of reference images. For this they proposed to split training images from representative data sets into small patches and learn a  Gaussian Mixture Model (GMM) modeling the distribution of extracted patches. To use the GMM in reconstruction they introduced the concept of expected patch likelihood (EPLL). They showed that the patch based GMM prior led to much improved image reconstructions for multiple inverse problems such as denoising, deblurring and inpainting. \cite{Bouman2016} later adapted the patch prior successfully for the reconstruction of radio data from the Event Horizon Telescope (EHT).

@@

    \subsection{Some Remarks on Deep Learning Methods}
    In the past decade Machine learning (ML) especially deep convolutional neural networks (CNN) became the standard for deblurring and super-resolution of natural images. However their success relies on the availability of massive sets of training data with typically low noise and corresponding ground truth to perform supervised learning. For a variety of reasons, which we will outline in the following, this approach is not ideal for astronomical x-ray and \gammaray data.

@@

    To account for the possibility of different datasets having a different number of pixels we divide by $N$. By summing over all observations we finally get the joint log-likelihood of measuring a set of counts images $\mathbf{D}$ given an unknown flux image $\mathbf{x}$:
    %
    \begin{equation}
        \label{eq:joint}
        \mathcal{L}\left( \mathbf{D} | \mathbf{x} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D_j} | \mathbf{x} \right)
    \end{equation}

@@

    \subsection{Patch Based Image Prior}
    In general it is difficult to capture global image statistics in a prior likelihood taking into account all pixel to pixel correlations. The size of the corresponding correlation matrix would grow with the number of pixels squared and thus quickly becomes computationally untraceable. However astronomical images typically only show local correlations between pixels. This is because the spatial scale of astrophysical processes in an image is always limited by the finite interaction time and the fact that the interaction cannot propagate faster than the speed of light (structure in the CMB might be an exception here). This means on smaller scales astronomical images often contain basic structures such as edges, corners, filaments or periodic patterns formed by local astrophysical processes. For this reason we adopt the idea of a patch based image prior approach as introduced by \cite{Zoran2011}. Given a pre-trained Gaussian Mixture Model (GMM) for the distribution of pixel values in patches, we can evaluate a patch based prior for the reconstructed image $\mathbf{x}$ and approximate the \aposteriori likelihood.

@@

    \cite{Zoran2011} found that the reconstruction quality could be improved by working with a grid of overlapping patches. Such a grid is illustrated in Figure~\ref{fig:patches}. However using a grid of overlapping patches makes the interpretation of the prior as likelihood more difficult, because pixels at the corners and edges of patches are counted multiple times. For this we introduce a weighting mask when evaluating the likelihood of the GMM. This mask weights the pixels of the patch by their overlap with the neighboring patch. Overlapping edges are assigned a weight of 1/2, while overlapping corners are assigned a weight of 1/4, when computing the likelihood for a patch as given by  Equation~\ref{eq:gmm}. Apart from this we keep the process of choosing the most likely component unmodified. However we note that in this case we neglected the correlation between neighbouring patches. As the choice of the GMM component per patch is discrete, a simultaneous gradient based optimization of all patches is not easily possible, because we cannot compute gradients across GMM components. This would require a continuous representation or interpolation between GMM components.

@@

    % \subsection{Choice and Tuning of Hyperparameters}
    % \subsubsection{Cross Validation}
    % Cross validation (CV) is an established method for model selection in statistics and machine learning to prevent a model from overfitting to the training data. To perform CV the available data is slit into a \enquote{training} and \enquote{validation} dataset. While optimizing the log-likelihood value is monitored  The use of multiple dataset allows to use cross-validation to prevent overfitting. One of the standard techniques in machine learning...

    % Cross validation is feasible once sufficient data is available, which is for example the 
    % case for deep \chandra~observations.

    % CV as stopping rule for RL was proposed by \cite{Reeves1995}...

    % \subsubsection{Other Approaches}

    % Self supervised Poisson loss? J-invariance has been shown to work, Grid Search? Calibration on similar data with ground truth?
    % Ground truth based calibration: record prior vs data likelihood on simulated data...

@@

    \subsubsection{Chandra SNRs}
    \label{sssec:chandra-snrs}. Because of the limited statistics we did not use any oversampling but kept the pixel Astrophysical processes forming the morphological structures are typically self-similar. Structures appear on multiple angular scales. This means the prior could even be learned from the data itself if sufficiently free of noise. For analysis of Galactic sources with complex morphology and typically radially symmetric features such as supernova remnants (SNRs) we learned a prior from images of the bright \chandra SNRs \textit{Cas A} and \textit{G292.0+01.8}. In total we extracted \todo{insert number} patches and learned a GMM with 128 components.

@@

    \begin{itemize}
        \item Results are available in \url{https://github.com/jolideco/jolideco-benchmarks}
        \item We tried the binary search tree as proposed in \cite{Parameswaran2018},
        but without success, most likely because the native \texttt{Pytorch} support
        for parallelization speeds up the standard GMM quite a bit...
        \item Could use a Neural Net for GMM component selection as proposed in \cite{Rosenbaum15}
        \item GPU support via Pytorch
    \end{itemize}

@@

    \begin{itemize}
        \item \textbf{Pylira}: first we use the LIRA method \citep{Esch2004} via the Pylira Python wrapper \citep{Donath2022}. We rely on the default prior definition and keep the background norm fixed. In total we run the the sampling process for 5000 iterations and use 500 iterations as a \endquote{burn-in} period to allow the MCMC chain to converge. Then we compute the mean of the posterior distribution ignoring the burn-in period.

@@

    On scenario A1 all methods show acceptable to excellent results for the isolated point sources in the upper right of the image. This demonstrates all methods take into account the PSF and background models correctly. The early stopped uniform prior still shows an extension for those cases, because the optimization process has not finished. Optimizing for 1000 iterations clearly improves the result and shows almost perfect point sources corresponding to the size of approximately one pixel. The \jolideco method with the Zoran-Weiss patch prior results in less sharp images, because the prior has been learned on natural images of landscapes and every day objects and scenes, which do not include point sources. The \jolideco method with the prior learned from GLEAM data, which also includes point sources, shows much improved results. However not as sharp as the results obtained with the uniform prior. In the lower left of the image the point sources become fainter and move closer, challenging all methods in succeeding to separate those as independent sources. The uniform prior optimized \enquote{to end} shows good results, still separating the sources. All other methods cannot fully resolve weak and close by point sources.

    For scenario B3 all methods recover the point sources well. However large difference can be observed in the reconstruction of the Gaussian shaped extended sources. As expected, the uniform prior assumption lets the emission decompose into many point sources, neglecting the spatial correlation between them. Pylira shows slightly improved results with the gaps between the bright emission peeks being more filled, however still lacks a convincing result of reconstructing the source as a single object. The \jolideco method with the patch priors shows the best results both visually as well as measured by the SSI. The GLEAM v0.1 patch prior reconstructs the sources almost perfectly, visually almost indistinguishable from the ground truth. The Zoran \& Weiss prior also shows good reconstruction of the extended sources, but less smooth and leaving a more visible extension in the surrounding point sources.

    Scenario C3 challenges the methods to reconstruct a weak extended source under four brighter point sources. While again all methods reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. Pylira shows a similar tendency but already improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features in the same scenario are decomposed into a \enquote{thread} of point sources by the uniform prior and LIRA method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that is possibly encoded in the training data.

@@

    \section{Application Examples}
    %
    \begin{figure*}
        \script{example-chandra.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               Result of the \jolideco method applied to \chandra data of the SNR \textit{1E~0102.2–7219}. The left large image shows the summed counts from 25 observations. The right large image  shows the result from the \jolideco deconvolution process. The flux is given as counts relative to the reference observation with ID=1308. The panels \textit{Zoom A}, \textit{Zoom B} and \textit{Zoom C} show a zoomed-in version of three example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
        \end{centering}
    \end{figure*}
    %
    \subsection{Deep \chandra Observation}
    \subsubsection{Image Example}
    As a first application example for real data we choose a set of short exposure observations of the SNR \textit{1E~0102.2–7219} by the \chandra telescope and combine them into a single deep observation using \jolideco and the patch prior learned from JWST data. In total we chose 25 observations\footnote{Same list of observation IDs as for \url{https://chandra.si.edu/photo/2009/e0102/}} taken with  the ACIS instrument and no grating.

    We reduced the data using \texttt{ciao} v4.15 and a custom \texttt{snakemake} workflow (see also Section~\ref{sec:reproducibility} on reproducibility). For the counts image we selected the energy range from \qty[mode = text]{0.5}{keV} to  \qty[mode = text]{7}{keV} and a pixel size of $0.25$ of the native pixel size of \chandra. This corresponds to an absolute  pixel size of $\approx$~\qty[mode = text]{0.125}{$\arcsec$/pixel} or a sampling of $\approx2$ pixels for the $38\%$ containment radius of Chandra's PSF. \chandra's instrument response varies with the observation offset and energy \citep{ChandraPOG2022}, so we simulated the PSF for each observation independently using the tool \texttt{marx} \citep{Davis2012} and the \texttt{simulate\_psf} script. As reference spectrum for the PSF simulation we assumed the model derived by \cite{Plucinsky2017}. We converted the spectral from XSPEC to sherpa script using the helper script \texttt{convert\_xspec\_script}. As reported in \cite{Xi2019} the effect of pileup, even for the brightest regions is small ($<5\%$), so we did not include the effect of pileup in our PSF simulations.

@@

    \textbf{Zoom B} shows the region around the central compact object of the SNR. While the emission is barely visible in the counts image it clearly sticks out above the background in the image reconstructed by \jolideco. The remaining visible extension in the point source is encoded in the JWST patch prior, which also shows a finite extension for point sources. We note that, as shown in Section~\ref{subsec:methods} for scenario A, the apparent size of the point source was smaller if the source was brighter. We estimated the position of point source by finding the brightest pixel in the map. We obtained $\alpha_{J2000}=01^h04^m02.75^s$ and $\delta_{J2000}=-72^d02^m00.19^s$ which is consistent with the value reported by \cite{Long2020}.

@@

    Checking the residual images per observation (not shown here), we noticed that the assumption of the time independence of the source as well the instrument response is not completely fulfilled in this case. We saw the detection efficiency of \chandra changed between the observations as well as that sub-structure in the remnant expands over time. Both effects are previously known and have been reported for example in \cite{Xi2019}. Despite the clear improvement in spatial resolution the proper motion of structures in the image might limit the quality of the result. The full analysis example is available at \url{https://github.com/jolideco/jolideco-chandra-examples}.

@@

    The profile computed from the images reconstructed with \jolideco has a much smaller noise level and the peak is much more pronounced. By fitting the profile with a univariate spline and finding its roots we measure the full width half maximum of the curves. We find $\lambda_J=0.56\pm0.03$ for the \jolideco profile and $\lambda_C=0.63\pm0.09$ for the profile computed on the counts image. Assuming that the PSF and the shape of peak of the profile are approximately Gaussian a we expect the following relation between the width of the counts profile, the PSF and the flux profile:

    \begin{equation}
        \lambda_{J} = \sqrt{\lambda_C^2 - \lambda_{\mathrm{PSF}}^2}
    \end{equation}

    With $\lambda_{\mathrm{PSF}} \approx 0.25 \mathrm{arcsec}$ \footnote{The 38\% containment radius of the Chandra PSF is $\sigma_{39}=2$ native pixels. This corresponds to  see e.g. \url{https://space.mit.edu/cxc/marx/tests/PSF.html}}
    we obtain ..., which is consistent with the values we measured for the profile. The profile computed from the images reconstructed with \jolideco show both a smaller width of the profile as well as a much smaller error.

@@

    The \texttt{fermitools} produce an energy dependent counts, exposure and PSF cube. To further reduce the data to images we use the \textit{Gammapy} package v1.0 \citep{GammapyZenodov1.0.1, Donath2023}. To compute an image of the expected background emission we rely on the latest Galactic diffuse emission model \texttt{gll\_iem\_v07.fits} and also include nearby sources from the 3FHL catalog \citep{Ajello2017}. We first fit the model components to the data using \textit{Gammapy}, excluding a circular region of radius \qty[mode = text]{1.02}{$\deg$} centered on Vela Junior. Finally we reduce this model to an image by summing over the energy dimension. To compute the exposure and PSF image we assume a power law with index $\Gamma=1.77$ \citep{Ajello2017} as a spectral model for Vela Junior and integrate over the energy dimension of the cube while weighting with the spectral model. We chose the GMM patch prior learned from \chandra SNRs described in the Section~\ref{sssec:chandra-snrs}. Because of the limited statistics we did not use any oversampling but keep the pixel size of the counts image.

@@

    Where $\hat{N}$ is the weighted sum of counts from a Gaussian smoothing kernel with $\sigma=5$ pixels. The result is shown in the most right panel of Fig.~\ref{fig:example-fermi-lat}. The image
    shows no systematic effects and a spatially homogeneous distribution of the residual noise. The resulting distribution of residual significances per pixel is centered at $\mu=-0.001$ with a width of $\sigma=0.991$, indicating a close agreement with a unit Gaussian distribution, as expected for a good model of the emission. More details of the example \fermi analysis are available at \url{https://github.com/jolideco/jolideco-fermi-examples}

      \begin{figure*}
        \script{example-fermi-lat.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-fermi-lat.pdf}
            \caption{
                Illustration of the result of the \jolideco method applied to \fermi data of the supernova remnant \textit{RX J0852.0-4622} or \textit{Vela Junior}. The left image shows the counts above \qty[mode = text]{10}{GeV}. The different event classes are stacked into a single image. The image in the center shows the flux reconstructed by the \jolideco method. The image on the right shows the residuals as computed by Equation~\ref{eq:approx-sigma} and smoothed with a \textit{Gaussian} kernel of width 5~pixels. More information on this analysis example can be found on \url{https://github.com/jolideco/jolideco-fermi-examples}.
            }
            \label{fig:example-fermi-lat}
        \end{centering}
    \end{figure*}
